---
title: "Cleaned_UP_Schemes"
author: "Rodoshee Das"
output:
  html_document: 
    code_folding: hide
    anchor_sections: FALSE
    toc: yes
    df_print: tibble
  pdf_document: default
  word_document: default
---
In this file we will merge and clean the datasheets to get a final dataset of all the classified educational expenditure schemes in Uttar Pradesh. 

[haq_unique_schemes](https://docs.google.com/spreadsheets/d/1lB9C6dZIJZErvjn4wwSXCK_qtyAOP9JY/edit?usp=sharing&ouid=107918389837609878151&rtpof=true&sd=true): containing the classification of the types of schemes & mode of benefit of transfer
[haq_data_renamed](https://docs.google.com/spreadsheets/d/1NxPfBjU9QD2y8f-EEAX9TMgOuK1T3err/edit?usp=sharing&ouid=107918389837609878151&rtpof=true&sd=true): containing all the schemes including the unclassified ones. 



```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(comment = NA)
```


# Load the two datasheets and libraries.

```{r}
library(readxl)
library(dplyr)
library(DT)
library(writexl) 
haq_data_renamed <- read_excel("C:/Users/Admin/Downloads/CivicDataLab/tasks/haq_data_renamed.xlsx")
View(haq_data_renamed)
haq_unique_schemes <- read_excel("C:/Users/Admin/Downloads/CivicDataLab/tasks/haq_unique_schemes.xlsx")
View(haq_unique_schemes)

```


# Load the dplyr package and merge the two datasheets based on the unique scheme codes

```{r}
library(dplyr)
haq_final <- merge(haq_data_renamed,haq_unique_schemes,by="scheme_code")

```

# remove the columns which aren't required for further analysis.

```{r}
haq_final$major_head <- NULL
haq_final$`sub-major_head`<- NULL
haq_final$...11 <- NULL
haq_final$`Classification of Schemes`<- NULL
haq_final$scheme_name.y <- NULL
haq_final$scheme_name.x <- NULL
haq_final$grant_number.y <- NULL
haq_final$grant_name.y <- NULL
```

#omit all the null values

```{r}
haq_final <- na.omit(haq_final)
```

# rearrange the columns for clarity of data.

```{r}
col_order <- c("scheme_code","Google Translation","Classification of the type of Scheme","Mode of Benefit Transfer","scheme_allotment","scheme_expenditure","scheme_utilisation","district_name","ddo_name","grant_number.x","fiscal_year","district_number","ddo_number","grant_name.x","division_name")

haq_final <- haq_final[,col_order]

```

#give the columns a proper name.
```{r}
colnames(haq_final)[which(names(haq_final) == "Google Translation")] <- "scheme_name"
colnames(haq_final)[which(names(haq_final) == "grant_number.x")] <- "grant_number"
colnames(haq_final)[which(names(haq_final) == "grant_name.x")] <- "grant_name"
datatable(haq_final)
```


# download the cleaned data frame in an excel datasheet.

```{r}
write_xlsx(haq_final,"C:/Users/Admin/Downloads/haq_final.xlsx")

```

