---
title: "EDA_UP_Schemes"
author: "Rodoshee Das"
date: "UP Fiscal Data_Scheme Analysis"
output:
  html_document:
      code_folding: hide
---

```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(comment = NA)
```

# Cleaned_UP_Schemes
In this file we will merge and clean the datasheets to get a final dataset of all the classified educational expenditure schemes in Uttar Pradesh. You can check this file in the repository.

[haq_unique_schemes](https://docs.google.com/spreadsheets/d/1lB9C6dZIJZErvjn4wwSXCK_qtyAOP9JY/edit?usp=sharing&ouid=107918389837609878151&rtpof=true&sd=true): containing the classification of the types of schemes & mode of benefit of transfer


[haq_data_renamed](https://docs.google.com/spreadsheets/d/1NxPfBjU9QD2y8f-EEAX9TMgOuK1T3err/edit?usp=sharing&ouid=107918389837609878151&rtpof=true&sd=true): containing all the schemes including the unclassified ones. 


# EDA_UP_Schemes

In this file, we did an exploratory data analysis of the dataset obtained from "Cleaned_UP_Schemes",i.e,haq_final.

Here, you can find interactive datatables and bar charts exploring:

1. Basic Structure of the dataset.
2. Levels of all the categorical variables.
3. Share of each of the classification of the type of scheme in the total expenditure.
4. Compare the expenditure of each district on different types of schemes.
5. Identify schemes with expenditure of zero.
6. Identify the grant type with highest amount of expenditure.
7. Identify the observations(/schemes) with the highest and lowest expenditure.
8. Identify the observations(/schemes) with the highest and lowest allotment.
9. Explore the share of each of the mode of benefit of transfer in the total expenditure.
10. Compare the expenditure of each district on modes of benefit of transfer.


# Load the dataset and libraries.

```{r}
library(readxl)
haq_final <- read_excel("C:/Users/Admin/Downloads/CivicDataLab/tasks/haq_final.xlsx")
View(haq_final)

library(dplyr)
library(ggplot2)
library(DT)
library(writexl)
```

## Check basic structure of the dataset.

```{r}
dim(haq_final)
str(haq_final)
```

## Start with the Exploratory Data Analysis (EDA).

```{r}

glimpse(haq_final)

unique(haq_final$scheme_code) # use unique() to get the levels of categorical variables
unique(haq_final$Scheme_Name)
unique(haq_final$`Classification of the type of Scheme`)
unique(haq_final$`Mode of Benefit Transfer`)
unique(haq_final$district_name)
unique(haq_final$ddo_name)
unique(haq_final$grant_number)
unique(haq_final$fiscal_year)
unique(haq_final$district_number)
unique(haq_final$ddo_number)
unique(haq_final$grant_name)
unique(haq_final$division_name)

```

# Explore the share of each of the classification of the type of scheme in the total expenditure.

```{r}

total_exp <- sum(haq_final$scheme_expenditure)

for(i in unique(haq_final$`Classification of the type of Scheme`)){
  df <- paste("Sch",i,sep = "_")
  assign(df, haq_final[haq_final$`Classification of the type of Scheme`==i,])
}

exp_boys <- sum(`Sch_Benefits boy students exclusively`$scheme_expenditure)
exp_boys/total_exp*100
exp_girls <- sum(`Sch_Benefits girl students exclusively`$scheme_expenditure)
exp_girls/total_exp*100
exp_both_directly <- sum(`Sch_Benefits both boy and girl students directly`$scheme_expenditure)
exp_both_directly/total_exp*100
exp_both_indirectly <- sum(`Sch_Benefits both boy and girl students indirectly`$scheme_expenditure)
exp_both_indirectly/total_exp*100

type_scheme <- c("Benefits boys only", "Benefits girls only", "Benefits both directly", "Benefits both indirectly")
share_scheme <- c(1.556087,0.2657823,97.26222,0.915911)
share_typesofschemes <- data.frame(type_scheme,share_scheme)
datatable(share_typesofschemes)



# produce a graph using ggplot for each of them.         
ggplot(data = share_typesofschemes, aes(x = type_scheme, y = share_scheme)) + geom_bar(aes(fill = share_scheme), stat = 'identity')


```


# compare the expenditure of each district on different types of schemes.

```{r}
onlygirls <- aggregate(x = `Sch_Benefits girl students exclusively`$scheme_expenditure,
                       by= list(`Sch_Benefits girl students exclusively`$district_number),
                       FUN=sum)
colnames(onlygirls)[which(names(onlygirls) == "Group.1")] <- "district_no"
colnames(onlygirls)[which(names(onlygirls) == "x")] <- "expenditure"
onlygirls <- onlygirls[order(onlygirls$expenditure),]
datatable(onlygirls)
```

```{r}
onlyboys <- aggregate(x = `Sch_Benefits boy students exclusively`$scheme_expenditure,
                      by= list(`Sch_Benefits boy students exclusively`$district_number),
                      FUN=sum)
colnames(onlyboys)[which(names(onlyboys) == "Group.1")] <- "district_no"
colnames(onlyboys)[which(names(onlyboys) == "x")] <- "expenditure"
onlyboys <- onlyboys[order(onlyboys$expenditure),]
datatable(onlyboys)
```


```{r}
bothdirectly <- aggregate(x = `Sch_Benefits both boy and girl students directly`$scheme_expenditure,
                          by= list(`Sch_Benefits both boy and girl students directly`$district_number),
                          FUN=sum)
colnames(bothdirectly)[which(names(bothdirectly) == "Group.1")] <- "district_no"
colnames(bothdirectly)[which(names(bothdirectly) == "x")] <- "expenditure"
bothdirectly <- bothdirectly[order(bothdirectly$expenditure),]
datatable(bothdirectly)
```


```{r}
bothindirectly <- aggregate(x = `Sch_Benefits both boy and girl students indirectly`$scheme_expenditure,
                            by= list(`Sch_Benefits both boy and girl students indirectly`$district_number),
                            FUN=sum)
colnames(bothindirectly)[which(names(bothindirectly) == "Group.1")] <- "district_no"
colnames(bothindirectly)[which(names(bothindirectly) == "x")] <- "expenditure"
bothindirectly <- bothindirectly[order(bothindirectly$expenditure),]
datatable(bothindirectly)

```

#create a dataframe for schemes whose expenditure is zero.

```{r}
exp_zero <- haq_final [(haq_final$scheme_expenditure == 0),]
datatable(exp_zero)

unique(exp_zero$scheme_utilisation) 
# can scheme utilization be anything other than 0 if expenditure is 0 !?    
```

# find out the grant type with highest amount of expenditure.

```{r}
for(i in unique(haq_final$grant_number)){
  df1 <- paste("Grant",i,sep = "_")
  assign(df1, haq_final[haq_final$grant_number==i,])
}

exp_grant14 <- sum(Grant_14$scheme_expenditure)
share_grant14 = exp_grant14/total_exp*100

exp_grant92 <- sum(Grant_92$scheme_expenditure)
share_grant92 = exp_grant92/total_exp*100

exp_grant22 <- sum(Grant_22$scheme_expenditure)
share_grant22 = exp_grant22/total_exp*100

exp_grant75 <- sum(Grant_75$scheme_expenditure)
share_grant75 = exp_grant75/total_exp*100

exp_grant72 <- sum(Grant_72$scheme_expenditure)
share_grant72 = exp_grant72/total_exp*100

exp_grant71 <- sum(Grant_71$scheme_expenditure)
share_grant71 = exp_grant71/total_exp*100

exp_grant47 <- sum(Grant_47$scheme_expenditure)
share_grant47 = exp_grant47/total_exp*100

exp_grant62 <- sum(Grant_62$scheme_expenditure)
share_grant62 = exp_grant62/total_exp*100

exp_grant83 <- sum(Grant_83$scheme_expenditure)
share_grant83 = exp_grant83/total_exp*100

exp_grant81 <- sum(Grant_81$scheme_expenditure)
share_grant81 = exp_grant81/total_exp*100

exp_grant48 <- sum(Grant_48$scheme_expenditure)
share_grant48 = exp_grant48/total_exp*100

exp_grant39 <- sum(Grant_39$scheme_expenditure)
share_grant39 = exp_grant39/total_exp*100

share_grants <- c(share_grant14,share_grant22,share_grant39,share_grant47,share_grant48,share_grant62,share_grant71,share_grant72,share_grant75,share_grant81,share_grant83,share_grant92)
name_grants <- c("14", "22","39", "47", "48", "62","71","72", "75", "81", "83","92")
grant_comparison <- data.frame(name_grants,share_grants)
datatable(grant_comparison)

ggplot(data = grant_comparison, aes(x = name_grants, y = share_grants)) + geom_bar(aes(fill = share_grants), stat = 'identity')
unique(Grant_71$grant_name)
unique(Grant_72$grant_name)

```


# identify the observations(/schemes) with the highest and lowest expenditure.

```{r}
max(haq_final$scheme_expenditure)  
highest_exp <- haq_final[(haq_final$scheme_expenditure == 14787739297), ]
datatable(highest_exp)
min(haq_final$scheme_expenditure)
lowest_exp <- haq_final[(haq_final$scheme_expenditure == 0), ]
datatable(lowest_exp)
unique(lowest_exp$grant_number)
```



```{r}
culturalaffair <- haq_final[(haq_final$grant_name =="CULTURAL AFFAIR"),]
datatable(culturalaffair)
unique(culturalaffair$grant_number) 
```

#So, 92(Cultural affairs)doesn't have any 0 expenditure scheme
#none of them directly benefits boys and girls, all of them are in Lucknow, only one in Gorakhpur
#maximum utilization - 100%
#this is the only grant which doesn't have zero scheme expenditure
#mode of benefit is infrastructure

# identify the observations(/schemes) with the highest and lowest allotment.
```{r}
max(haq_final$scheme_allotment)       
highest_allot <- haq_final[(haq_final$scheme_allotment == 14787739297), ]
datatable(highest_allot)

min(haq_final$scheme_allotment)
lowest_allot <- haq_final[(haq_final$scheme_allotment < 0), ]
datatable(lowest_allot)

#negative allotment amount!?


```

# Explore the share of each of the mode of benefit of transfer in the total expenditure.
```{r}
total_exp <- sum(haq_final$scheme_expenditure)

for(i in unique(haq_final$`Mode of Benefit Transfer`)){
  df <- paste(i)
  assign(df, haq_final[haq_final$`Mode of Benefit Transfer`==i,])
}

cash_transfer_exp <- sum(`Direct Cash Transfer to students`$scheme_expenditure)
cash_transfer = cash_transfer_exp/total_exp*100
In_Kind_exp <- sum(`In-kind service delivery`$scheme_expenditure)
In_Kind = In_Kind_exp/total_exp*100
Infra_exp <- sum(Infrastructure$scheme_expenditure)
Infra = Infra_exp/total_exp*100
Inst_Grants_exp <- sum(`Institutional Grants`$scheme_expenditure)
Inst_Grants = Inst_Grants_exp/total_exp*100
PA_exp <- sum(`Payments & Awards`$scheme_expenditure)
Payments_Awards = PA_exp/total_exp*100
others_exp <- sum(Others$scheme_expenditure)
Oth = others_exp/total_exp*100

type_mode <- c("Cash Grants","In-kind Delivery","Infrastructure","Institutional Grants","Payments and Awards","Others")
share_mode <- c(cash_transfer,In_Kind,Infra,Inst_Grants,Payments_Awards,Oth)
share_modesoftransfer <- data.frame(type_mode,share_mode)
datatable(share_modesoftransfer)

                                          
ggplot(data = share_modesoftransfer, aes(x = type_mode, y = share_mode)) + geom_bar(aes(fill = share_mode), stat = 'identity')

```

# compare the expenditure of each district on modes of benefit of transfer.


```{r}
direct_cash <- aggregate(x = `Direct Cash Transfer to students`$scheme_expenditure,
                       by= list(`Direct Cash Transfer to students`$district_number),
                       FUN=sum)
colnames(direct_cash)[which(names(direct_cash) == "Group.1")] <- "district_no"
colnames(direct_cash)[which(names(direct_cash) == "x")] <- "expenditure"
direct_cash <- direct_cash[order(direct_cash$expenditure),]
datatable(direct_cash)
```


```{r}
in_kind <- aggregate(x = `In-kind service delivery`$scheme_expenditure,
                         by= list(`In-kind service delivery`$district_number),
                         FUN=sum)
colnames(in_kind)[which(names(in_kind) == "Group.1")] <- "district_no"
colnames(in_kind)[which(names(in_kind) == "x")] <- "expenditure"
in_kind <- in_kind[order(in_kind$expenditure),]
datatable(in_kind)
```


```{r}
Infra <- aggregate(x = Infrastructure$scheme_expenditure,
                     by= list(Infrastructure$district_number),
                     FUN=sum)
colnames(Infra)[which(names(Infra) == "Group.1")] <- "district_no"
colnames(Infra)[which(names(Infra) == "x")] <- "expenditure"
Infra <- Infra[order(Infra$expenditure),]
datatable(Infra)
```


```{r}
Inst.Grants <- aggregate(x = `Institutional Grants`$scheme_expenditure,
                     by= list(`Institutional Grants`$district_number),
                     FUN=sum)
colnames(Inst.Grants)[which(names(Inst.Grants) == "Group.1")] <- "district_no"
colnames(Inst.Grants)[which(names(Inst.Grants) == "x")] <- "expenditure"
Inst.Grants <- Inst.Grants[order(Inst.Grants$expenditure),]
datatable(Inst.Grants)
```


```{r}
OtherModes <- aggregate(x = Others$scheme_expenditure,
                     by= list(Others$district_number),
                     FUN=sum)
colnames(OtherModes)[which(names(OtherModes) == "Group.1")] <- "district_no"
colnames(OtherModes)[which(names(OtherModes) == "x")] <- "expenditure"
OtherModes <- OtherModes[order(OtherModes$expenditure),]
datatable(OtherModes)
```


```{r}
Payments_Awards <- aggregate(x = `In-kind service delivery`$scheme_expenditure,
                     by= list(`In-kind service delivery`$district_number),
                     FUN=sum)
colnames(Payments_Awards)[which(names(Payments_Awards) == "Group.1")] <- "district_no"
colnames(Payments_Awards)[which(names(Payments_Awards) == "x")] <- "expenditure"
Payments_Awards <- Payments_Awards[order(Payments_Awards$expenditure),]
datatable(Payments_Awards)
```



